Parallel algorithm for solving multicriterial optimization problems using elements of machine learning

### 2025-06-18 2025/RSCD_MCO/RSD-2025.tex
fix English in the comments

In the framework of the developed approach \cite{Konnov2025}, the second option with scalarization of criteria is used. To build a good estimate of the Pareto set, a family of problems
\begin{equation}
\label{eq:04}
\min {\varphi_j (y)}=F(\lambda_j,y),F(\lambda_j,y) \in \{F(\lambda_1,y),F(\lambda_2,y),\dots,F(\lambda_\tau,y)\},y\in D,
\end{equation}
is solved,  where $F(\lambda_j,y)$, $1 \leq j \leq \tau$, is a multiextremal function, obtained on the base of scalarization of criteria $f_i(y)$, $1 \leq i \leq s$. Here $\lambda_j=\{\lambda_{i,j}\}$, $1 \leq i \leq s$, $1 \leq j \leq \tau$, are significance indicators for each criterion, s is the number of criteria, $\tau$ is the number of scalar optimization problems. There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}. The performed computational experiments used minimax convolution of criteria
\begin{equation}
\label{eq:05}
F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D, \sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
\end{equation}

@@

Computational experiments were carried out on the nodes of the Lobachevsky supercomputer of Nizhny Novgorod State University with two processors Intel Xeon Silver 4310T 2.3 GHz, 64 GB RAM. Machine learning libraries SciKit-Learn 0.24.2 \cite{scikit-learn} and PyTorch 2.5.1 \cite{PyTorch} were used in experiments. The algorithm proposed in the work was implemented in Python using the framework iOpt \cite{ioptmco}.

>>>>

In the framework of the developed approach \cite{Konnov2025}, the second option with scalarization of criteria is used. To build a good estimate of the Pareto set, a family of problems
\begin{equation}
\label{eq:04}
\min {\varphi_j (y)}=F(\lambda_j,y),F(\lambda_j,y) \in \{F(\lambda_1,y),F(\lambda_2,y),\dots,F(\lambda_\tau,y)\},y\in D,
\end{equation}
is solved,  where $F(\lambda_j,y)$, $1 \leq j \leq \tau$, is a multiextremal function, obtained on the base of scalarization of criteria $f_i(y)$, $1 \leq i \leq s$. Here $\lambda_j=\{\lambda_{i,j}\}$, $1 \leq i \leq s$, $1 \leq j \leq \tau$, are significance indicators for each criterion, s is the number of criteria, $\tau$ is the number of scalar optimization problems. There are different ways to build functions $F(\lambda_j,y)$ \cite{Miettinen1999,Konnov2025,Marler2004}.  
In the proposed algorithm we used fixed value $\tau = 50$. The values $\lambda_j$, $1 \leq j \leq \tau$, are selected uniformly within the set of values satisfying the condition (\ref{eq:051})
\begin{equation}
\label{eq:051}
\sum_{i=1}^s {\lambda_{i,j}} = 1, 0 \leq \lambda_{i,j} \leq 1.
\end{equation}

The performed computational experiments used minimax convolution of criteria

\begin{equation}
\label{eq:05}
F(\lambda_j, y) = \max_{1 \leq i \leq s} \left(\lambda_{i,j} f_i (y)\right), y \in D.
\end{equation}

@@

Computational experiments were carried out on the nodes of the Lobachevsky supercomputer of Nizhny Novgorod State University with two processors Intel Xeon Silver 4310T 2.3 GHz, 64 GB RAM. Machine learning libraries SciKit-Learn 0.24.2 \cite{scikit-learn} and PyTorch 2.5.1 \cite{PyTorch} were used in experiments. The algorithm proposed in the work was implemented in Python using the framework iOpt \cite{ioptmco}. 
% Результаты сравнения последовательной версии предлагаемого алгоритма с современными метаэвристическими решателями представлены в нашей статье \cite{Konnov2025}. Текущая работа сосредоточена на анализе эффективности параллельной версии алгоритма.
The results of comparing the sequential version of the proposed algorithm with state-of-the-art metaheuristic solvers are presented in our paper \cite{Konnov2025}. The current work focuses on analyzing the efficiency of the parallel version of the algorithm.
